Stack Safety is a Security Property

### 2021-04-10 tex/main.tex
Rewrite abstract; work on intro

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\apt{maybe: ``Stack safety is a testable security property'' ?}a
  What is stack safety? The phrase has been used to describe the protections
  offered by a variety of mechanisms for guarding against
  attacks that corrupt the call stacks of running programs, including micro-policies
  implemented using metadata tagging mechanisms such as PIPE. But for specification
  these implementations largely appeal either to intuitive notions of the behavior
  of a stack, or to specific examples of bad behavior.
  %
  We propose a formal definition phrased in the technical terminology of language-based
  security---as a combination of an integrity property (a caller's stack frame is
  invariant over calls) and a confidentiality property (callees lack knowledge of the
  contexts of their calls). These combine with a straightforward well-bracketedness
  condition on the program's control flow that ensures that callers and callees remain
  meaningfully distinct.
  %
  While our formalism is informed by the PIPE implementation, the primary properties reflect
  general principles of stack safety. We define additional properties tailored for specific
  contexts: a stronger property amenable to efficient testing and a weaker one that is
  approximated by more efficient, lazy policies.

  \ifaftersubmission\bcp{should we mention stktokens here?}
  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
  \sna{On the same note, cite Nick and Andre?}
  \fi
\end{abstract}

@@

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
% 
The implementation and configuration of these mechanisms can be complex, 
and often involves making compromises that trade off protection for efficiency.  
Thus it would be very useful to have precise formal specifications for stack safety
that give us a basis for comparing the security guarantees claimed by different enforcement
techniques, and for verifying that these claims are met by particular implementations.

@@

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
\apt{more to say?}

@@

In summary, we offer the following contributions:
\begin{itemize}
\item 
  We harness the language-based security concepts of integrity and confidentiality 
  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We apply a property-based randomized testing framework to provide evidence that 
  this stepwise property is enforced by the Depth Isolation policy
  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
\item 
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property, and argue that it accurately characterizes 
  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18},
  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item 
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
  to leave it out of the title but hit it strongly in the abstract.}%
What is stack safety? The phrase is associated with a variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms mostly lack rigorous specifications, appealing
either to intuitive expectations of how good stack-using programs should
behave or to specific examples of bad behavior.

We propose a formal criterion for stack safety, phrased in terms of
language-based security: a combination of an integrity property (``each
caller's stack frame is invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's stack frame''). These combine with a straightforward
well-bracketedness condition on the program's control flow that ensures that
callers and callees remain meaningfully distinct.\bcp{I am still not clear
  how well-bracketedness fits into the story.  (For me, it is the bracketing
  that {\em defines} where calls and returns happen, so there is something
  tautological about saying that `` well-bracketedness condition ensures
  that callers and callees remain meaningfully distinct''!)}

Our criterion is inspired by hardware tagging mechanisms for stack
protection, in particular the stack-safety ``micro-policies''
of~\citet{DBLP:conf/sp/RoesslerD18}, but its core reflects general
principles of stack safety\bcp{I wish we did not have to be so apologetic
  about being tied to micro-policies (and it is still not clear to me that
  we have to be... do we have a clear statement someplace of why our
  definition would not apply to CFI, for example? This seems like a critical
  discussion...)}. We define additional properties tailored for specific
contexts: a stronger property amenable to efficient testing and a weaker one
that is approximated by more efficient, lazy policies.

\bcp{... and now a sentence on testing!}

  \ifaftersubmission\bcp{should we mention stktokens here?}
  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
  \bcp{Some people say not to, but I think their argument is weak: I've
    added Nick/Andre above.  But I'm not sure we want to get into stkTokens
    in the abstract.}
  \fi
\end{abstract}

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, with consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}
\fi
%
The implementation and configuration of these  stack-safety mechanisms can be complex,
and often involves compromises trading protection for efficiency.
Thus it would be useful to have a precise formal specification for stack
safety, both as a basis for comparing the security guarantees claimed by different enforcement
techniques and for verifying that these claims are met by particular implementations.

@@

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
  significant contribution, and partly to give the reader some context for
  why this is part of our story.  E.g., ``One nice thing we can do with our
  spec is to use it to evaluate the claims of existing or future proposals
  for enforcement mechanisms.  One relatively inexpensive way to carry out
  such an evaluation is to see whether property-based testing against this
  spec can distinguish correct implementations of the mechanism from buggy
  ones...''}

@@

In summary, we offer the following contributions:
\begin{itemize}
\item
  We adapt formulations of integrity and confidentiality from
  language-based security
  to formalize a parameterized {\em stepwise stack-safety property} and argue that
  it captures the intuitive concepts of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We apply a property-based randomized testing framework to provide evidence that
  this stepwise property is enforced by the Depth Isolation policy
  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
\item
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property and argue that it accurately characterizes
  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18}
  enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and simple
  coroutines~(\cref{sec:??}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}
